Educational attainment and intergenerational mobility: A polygenic score analysis

### 2025-01-07 posts/RustichiniIaconoLeeMcGue/RILM_PGSEdu_JPE2023.qmd
initial commit

* US Minnesota Twin Study Sample
* Build a genetic model (with better genetics foundation than **ad hoc** econ standard model)
* Intergenerational elaciticity of income is predicted and estimated to be larger in genetic models
* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect
* Parental PGS &nrArr; educational attainment, after controling for parental education and incomes

@@

@@

@@

@@

::::{.column width="45%"}
![recursive SEM](Tab3.jpg)
::::

::::{.column width="55%"}
Dizygotic twins (fraternal twins)

@@

::::{.column width="75%"}
* PGS of parents are not related to outcomes once IQ, soft skills index, parental education and family income are added
* Most parental impacts on education are through incomes and parental education
* No other indirect pathway remain from parental PGS, after parental education and incomes are addressed
   *  No genetic nurture
::::

@@

* Multiple testing気にせず

>>>>

* US Minnesota Twin Study Sample
* Build a genetic model of human capital investments with genotype transmission (with better genetics foundation than *ad hoc* standard econ models)
* Intergenerational elaciticity of income is predicted to be larger in genetic models
* Empirical pathway: Through parental education and incomes, parental PGS has no direct effect on child human capital
* Child PGS is relevant
* Genetic nurture (introverted genotype produces an environment to grow introverted child, independently of income and education) is not found

@@

:::{style="text-align: center;"}
# Model 
:::

@@

::: {style="font-size: 80%;"}
Family $i$ parents' problem with twins $j=1, 2$ of observable skills $\theta^{i}_{j}$
$$
\begin{aligned}
\max_{\{E^{i}, I^{i}_{1}, I^{i}_{2}\}} \;\;\;&\E\left[(1-\delta)\ln E^{i}+\delta\left(y^{i}_{1}+y^{i}_{2}\right)\left|\theta^{i}_{1}, \theta^{i}_{2}\right.\right]\\
\st \;\;\; & E^{i}+I^{i}_{1}+I^{i}_{2}=e^{y^{i}}
\end{aligned}
\tag{6}
$$
$$
\begin{align}
\scriptsize{\mbox{Human capital production function}} && 
h^{i}_{j}&=\alpha_{I}\ln I^{i}_{j}+\alpha_{\theta}\theta^{i}_{j}+\epsilon^{\mathrm{h}i}_{j} && \phantom{mmmmmmmmmmmm} \tag{8}\label{eq8}\\
\scriptsize{\mbox{Child income}} && 
y^{i}_{j}&=\alpha_{\mathrm{h}}h^{i}_{j}+\epsilon^{\mathrm{y}i}_{j} &&\tag{9}\label{eq9}
\end{align}
$$
Solution
$$
I^{i}=\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}e^{y^{i}}
\tag{11}
$$
Optimised child income
$$
y^{i}_{j}=\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\ln 
\frac{\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}{1-\delta+2\delta\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}}
+\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta^{i}_{j}+\alpha_{\mathrm{h}}\epsilon^{\mathrm{h}i}_{j}+\epsilon^{\mathrm{y}i}_{j}
\tag{12}\label{eq12}
$$ 
:::

Matching of parents

* Assume independence of genotype conditional on observable characteristics [incomes, skills $w(g)$]
   * Matches are made on observables, not on genotype
   * Matches are random within obserbable strata *"worth" classes*=$w_{\mathrm{y}}y+w(g)$]
      * Genotype matching is random within a stratum
   * Matches are nonrandom at the population due to population stratification based on observables

Genetic transmission: $\Delta$(father genotype) × $\Delta$(mother genotype) $\to$ $\Delta$(child genotype)

The appendix D of paper shows an invariant mapping exists with properties shown in Theorem 3.1

Skills production function (genotype enters here, hence the parental genotypes)
$$
\theta^{i}_{j}=w\left(g^{i}_{j}\right)+\pi y^{i}+\bfPi\bfx^{i}_{j}+F^{i}+\epsilon^{\theta i}_{j}
\tag{4}\label{eq4}
$$
Estimating equations: Substitute \eqref{4} into  \eqref{12} gives child income decomposition, and \eqref{11} into  \eqref{8} gives human capital decomposition
$$
\begin{align}
y^{i}_{j}
&=
\alpha_{\theta}\alpha_{\mathrm{h}}w\left(g^{i}_{j}\right)
+\left(\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}+\alpha_{\theta}\alpha_{\mathrm{h}}\pi\right)y^{i}
+\alpha_{\theta}\alpha_{\mathrm{h}}F^{i}\\
&\hspace{1em}
+\alpha_{\theta}\alpha_{\mathrm{h}}\bfPi\bfx^{i}_{j}
+\alpha_{\theta}\alpha_{\mathrm{h}}\epsilon^{\theta i}_{j}
+\alpha_{\mathrm{h}}\epsilon^{\mathrm{h} i}_{j}
+\epsilon^{\mathrm{y}i}_{j}\tag{21}\label{eq21}\\
h^{i}_{j}
&=
\alpha_{\theta}w\left(g^{i}_{j}\right)
+\left(\alpha_{\mathrm{I}}+\alpha_{\theta}\pi\right)y^{i}
+\alpha_{\theta}F^{i}\\
&\hspace{1em}
+\alpha_{\theta}\bfPi\bfx^{i}_{j}
+\alpha_{\theta}\epsilon^{\theta i}_{j}
+\epsilon^{\mathrm{y}i}_{j}\tag{23}\label{eq23}
\end{align}
$$

Can show IGE is smaller in "standard model" than genetic model

::: {style="font-size: 80%;"}
@BeckerTomes1979
$$
\begin{align}
\theta_{t+1}
&=
\eta \theta_{t}+\epsilon^{\theta}_{t+1}\tag{26}\label{eq26}\\
y^{i}_{t+1}
&=
\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}y^{i}_{t}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{25}\label{eq25}
\end{align}
$$
Genetic model of this paper assumes
$$
\begin{align}
\theta
&=
w^{\theta}_{\mathrm{f}}\theta_{\mathrm{f}}+w^{\theta}_{\mathrm{m}}\theta_{\mathrm{m}}\tag{A4}\label{eqA4}\\
y^{i}
&=
w^{\mathrm{y}}_{\mathrm{f}}y^{i}_{\mathrm{f}}+w^{\mathrm{y}}_{\mathrm{m}}y^{i}_{\mathrm{m}}\tag{A3}\label{eqA3}
\end{align}
$$
Plugging these give
$$
\begin{align}
\theta_{t+1}
&=
\eta \sum_{s=m,f}w^{\theta}_{s}\theta_{st}+\epsilon^{\theta}_{t+1}\tag{33}\label{eq33}\\
y^{i}_{t+1}
&=
\alpha_{\mathrm{I}}\alpha_{\mathrm{h}}\sum_{s=m,f}w^{\mathrm{y}}_{s}y^{i}_{st}+\alpha_{\theta}\alpha_{\mathrm{h}}\theta_{t+1}+\epsilon^{\mathrm{y}i}_{t+1}\tag{32}\label{eq32}
\end{align}
$$
:::

:::{style="text-align: center;"}
# Estimation strategy 
:::

Structural equation model (SEM)

* Structural equations
* Measurement (error) equations

Identification conditions of structural parameters remain the same as in simultaneous equation models

In this paper:

* identification for only recursive model is discussed (Tab 2)
* use of IVs is not discussed (Tab 3)

@@

Identification ideas

1. Dizygotic twins [Table 1, 2]
   * Share birth days, environment
   * Differ in genotypes
   * Ideal to see impacts of genotypes on outcomes

Identification ideas

2. SEM
   * Recursive system (37)-(42) are identified as usual [Table 2]
   * Structural parameters of endogenous variables are recovered from \eqref{eq43}-\eqref{eq45} (how?) [Table 3]
      * Given PGS of child, PGS of parents should be redundant, so long as they do not affect environment (gene-environment correlation, or "passive rGE")

::: {style="font-size: 80%;"}
$$
\begin{align}
e^{i}_{\mathrm{h}}
&=
\alpha_{e_{\mathrm{h}}}+
\gamma^{\mathrm{pGS}_{m}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{e_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{e_{\mathrm{h}}}\tag{43}\label{eq43}\\
y^{i}_{\mathrm{h}}
&=
\alpha_{y_{\mathrm{h}}}+\gamma^{\mathrm{pGS}_{m}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{m}+\gamma^{\mathrm{pGS}_{f}}_{y_{\mathrm{h}}}\mathrm{pGS}^{i}_{f}+\xi_{y_{\mathrm{h}}}\tag{44}\label{eq44}\\
e^{i}_{j}
&=
\alpha_{e}+\gamma^{e_{\mathrm{h}}}_{\mathrm{e}}e_{\mathrm{h}}^{i}
+\gamma^{y_{\mathrm{h}}}_{\mathrm{e}}y_{\mathrm{h}}^{i}
+\gamma^{\mathrm{pGS}}_{\mathrm{e}}\mathrm{pGS}^{i}_{j}
+\gamma^{\mathrm{pGS}_{m}}_{\mathrm{e}}\mathrm{pGS}^{i}_{m}
+\gamma^{\mathrm{pGS}_{f}}_{\mathrm{e}}\mathrm{pGS}^{i}_{f}
+\xi_{\mathrm{e}}\tag{45}\label{eq45}
\end{align}
$$
:::

@@

::::{.column width="45%"}
![full SEM](Tab3.jpg)
::::

::::{.column width="55%"}
Dizygotic (fraternal) + monozygotic (identical) twins

@@

::::{.column width="75%"}
* PGS of parents are not related to outcomes once IQ, soft skills index, parental education and family income are added
* Most parental impacts on education are through incomes and parental education
* No other indirect pathway remain from parental PGS, after parental education and incomes are addressed
   *  No passive rGE other than parental education and incomes (genetic nurture) is found
::::

@@

* 咀嚼できていません
* 人的資本投資を内生化したsteady state genetic transmission mappingの存在を証明
* それと実証部分の結びつきが弱い(理解できていない)
* 親の教育と所得を考慮すると、遺伝情報は子の教育に影響しないことを実証的に示したのは斬新
* SEMはidentification assumptionが不明(recursiveモデル以外)、IVを使っていない?
* Multiple testing気にせず
